### 2025-06-29 manuscript/part2.tex
module decomp done

Modules  .... register services, expose interfaces, auto generated service added to DI container and custom service added to DI container

Although the crucial segments are located inside \code{/src}, files of interest ... Szntax config

syntax TextMate Monarch(ignored as no web solution or sandboxing is drawn from only vs code -> textmate)

Lastly, situated at top-level in the \verb|src| directory are three files: \verb|setup....ts| \verb|setup....ts| \verb|setup....ts|, which are ...
by ... to ...

Lastly, situated at top-level in the \code{src/} directory are three files: `setupCommon.ts`, `setupClassic.ts`, and `setupExtended.ts`, which serve as entry points for configuring your language environment. Each exports a factory function used by your VS Code extension and CLI generators to assemble and wire up the Langium services with the appropriate feature set — `setupCommon` for shared core services, `setupClassic` for the standard LSP-focused configuration, and `setupExtended` for an enhanced setup that includes advanced validation, formatting, and other extended capabilities.

\begin{description}
  \item[\texttt{cli/cli-util.ts}] Shared helper routines for the CLI (file I/O, path resolution, error reporting).
  \item[\texttt{cli/generator.ts}] Yeoman-based scaffolder that bootstraps a new Langium DSL project with templates and example code.
  \item[\texttt{cli/main.ts}] Entry point for the CLI: parses arguments and dispatches subcommands (generate, validate, serve-lsp).
  \item[\texttt{directory.txt}] Snapshot of the directory structure, typically generated by \texttt{tree}, for documentation.
  \item[\texttt{extension/main.ts}] VS Code extension bootstrap: launches the LSP server, sets up the client, handles activation/deactivation.
  \item[\texttt{language/generated/ast.ts}] Auto-generated interfaces for each AST node, defining the DSL’s abstract syntax.
  \item[\texttt{language/generated/grammar.ts}] Auto-generated Chevrotain grammar: token definitions, regex patterns, and rules.
  \item[\texttt{language/generated/module.ts}] DI module binding parser, lexer, CST visitor, and other services into Langium’s registry.
  \item[\texttt{language/hello-world.langium}] The Langium grammar file: terminal and parser rule definitions in Langium DSL.
  \item[\texttt{language/hello-world-module.ts}] Extension of the generated module for custom services (linkers, formatters, completions).
  \item[\texttt{language/hello-world-validator.ts}] Hand-written semantic validation rules (e.g., name collisions, type checks).
  \item[\texttt{language/main-browser.ts}] Browser entry point for embedding the language (e.g., in Monaco), using WebSocket/Worker transport.
  \item[\texttt{language/main.ts}] Node-based language-server entry point: configures FS adapter, instantiates services, starts LSP.
  \item[\texttt{setupCommon.ts}] Registers services common to both Node and browser (grammar access, CST→AST visitor, linking, validation).
  \item[\texttt{setupExtended.ts}] Builds on \texttt{setupCommon} to add IDE features (formatting, folding, hover, semantic tokens).
  \item[\texttt{setupClassic.ts}] Classic Node setup: invokes \texttt{setupExtended}, plugs in Node FS and LSP server, exports startup function.
\end{description}

>>>>

The generated module is created automatically from the grammar definition and provides two DI fragments: the GeneratedSharedModule, which registers shared-core services such as AST reflection, and the GeneratedModule, which configures the parser, Chevrotain grammar factory, and language metadata. Conversely, the main module is handwritten: it imports Langium’s default core and shared modules, injects the generated fragments, declares custom services (validators, scope providers, caches), invokes the \texttt{inject} function to assemble the complete service container, and exposes a \texttt{createServices} factory.

Although the core logic resides in `/src`, files of interest at the project root include the standard npm `package.json` and two Langium-specific configs:
\code{language-configuration.json}, which governs editor behaviors such as comment delimiters, bracket pairs and auto-closing rules, and
\code{langium-config.json}, which declares settings like the language ID, grammar entry point, target file extensions and maximum lookahead.
In particular, the `syntaxes` entry drives the generation of a `.monarch.ts` module (for Monaco-based editors) and a `.tmLanguage.json` file (for TextMate-based engines).
The TextMate grammar maps regex patterns to semantic scopes that underpin VS Code's coloring, bracket
matching and code folding, and is the one applied for the extension. Customizing these definition files allows developers to implement nuanced, context-sensitive highlighting—such as semantic theming
and advanced pattern recognition—that cannot be exhaustively defined by grammar rules alone \cite{EDKarlsson2022Discussion604}.
